---
title: "HW1"
author: "Colin Wick"
date: "1/22/2021"
output:
  html_document: default
  md_document:
    variant: markdown_github
---

## Question 1

```{r,warning=FALSE,cache=TRUE}
library(tidyverse)
gas <- read.csv("../../../../../GitHub/ECO395M/data/GasPrices.csv")

gas %>%
  group_by(Competitors) %>%
  summarize(avg_price = mean(Price))

gas %>%
  group_by(Competitors) %>%
  ggplot()+
  geom_boxplot(aes(y=Price,x=Competitors))

```

There is modest evidence that gas stations within sight of competitors charge lower prices. They also tend to have less variance in prices.

```{r,warning=FALSE,cache=TRUE}
gas %>%
  ggplot()+
  geom_point(aes(y=Price,x=Income))+
  geom_smooth(method=lm,aes(y=Price,x=Income))
```

Strong evidence suggesting richer areas have higher prices.

```{r,warning=FALSE,cache=TRUE}
gas %>%
  group_by(Brand) %>%
  summarize(avg_price = mean(Price)) %>%
  ggplot()+
  geom_col(aes(x=Brand,y=avg_price))
```

No evidence that any brand charges particularly higher prices.

```{r,warning=FALSE,cache=TRUE}
gas %>%
  ggplot()+
  geom_histogram(aes(Price))+
  facet_grid(rows = vars(IntersectionStoplight))
```

There is not enough data to make a clear pronouncement, but gas stations next to both a stoplight and intersection appear to charge slightly more.

```{r,warning=FALSE,cache=TRUE}
gas %>%
  group_by(Highway) %>%
  ggplot()+
  geom_violin(aes(y=Price,x=Highway))

gas %>%
  group_by(Highway) %>%
  ggplot()+
  geom_histogram(aes(x=Price))+
  facet_grid(rows=vars(Highway))

```

Stations near the highway charge more than those far from the highway.

## Question 2

```{r,warning=FALSE,cache=TRUE}
bike <- read.csv("../../../../../GitHub/ECO395M/data/bikeshare.csv")

bike %>%
  ggplot()+
  geom_boxplot(aes(hr,total,group=hr))+
  geom_line(aes(hr,total),stat = "summary", fun.y = "mean",size=2,alpha=.2,color="blue")
```

Average bike rentals per hour of the day. Average line is in blue, while boxplots show the shape of the variance.

```{r,warning=FALSE,cache=TRUE}
bike %>%
  ggplot()+
  geom_boxplot(aes(hr,total,group=hr))+
  geom_line(aes(hr,total),stat = "summary", fun.y = "mean",size=2,alpha=.2,color="blue")+
  facet_grid(rows = vars(workingday))
```

Faceting on whether it is a working day, it's clear that on workdays there is higher demand around commuting hours while on weekends the mid-afternoon tends to be peak demand.

```{r,warning=FALSE,cache=TRUE}
workday_labs <- c("NW", "W")
names(workday_labs) <- c("0", "1")

bike %>%
  filter(hr=="8") %>%
  mutate(weathersit = as.factor(weathersit))%>%
  ggplot() +
  geom_bar(aes(x=weathersit,y=total,fill=weathersit),stat = "summary", fun.y = "mean")+
  facet_grid(. ~ workingday, labeller = labeller(workingday = workday_labs))+
  scale_x_discrete(labels = c("Clear","Mist/Cloud","Rain"))+
  theme(axis.title.x = element_blank(),legend.position = "none")+
  ylab(label = "average rides")
```

"NW" represents nonworking days, where we see a significant difference if it is raining versus non-raining days.

# Question 3
```{r,warning=FALSE,cache=TRUE}

air <- read.csv("../../../../../GitHub/ECO395M/data/ABIA.csv")

air$Date <- as.Date(paste(air$Year,air$Month,air$DayofMonth,sep="-"),format="%Y-%m-%d")

air_taxi <- air %>%
  select(Origin,WeatherDelay,TaxiIn,TaxiOut,FlightNum,Month,DepTime,ArrTime) %>%
  mutate(WeatherDelay = ifelse(is.na(air$WeatherDelay),0,air$WeatherDelay),
         unique_id = paste(air$Date,FlightNum,sep="-"),
         to_from = ifelse(Origin=="AUS","Outbound","Inbound")) %>%
  pivot_longer(cols=c("TaxiIn","TaxiOut"),names_to="Taxi",values_to="TaxiTime")
```



```{r,warning=FALSE,cache=TRUE}
air %>%
  filter() %>%
  ggplot()+
  geom_bin2d(aes(TaxiIn,TaxiOut))+
  geom_smooth(aes(TaxiIn,TaxiOut),method = "lm")
```

Plotting TaxiIn and TaxiOut density against each other and adding an OLS line to see relationships. We see a weak but noticeable correlation between TaxiIn and TaxiOut times. There is no reasonable expectation that these two would have a causal relationship with each other. What causes the variance in TaxiIn and TaxiOut?

```{r,warning=FALSE,cache=TRUE}
air_taxi %>%
  ggplot()+
  geom_histogram(aes(TaxiTime),bins = 50)+
  facet_grid(Taxi~to_from)+
  xlim(0,50)
```

This breakdown shows histograms of taxi times based on whether the taxi was leaving or arriving at the airport and whether the flight was inbound or outbound.

TaxiIn + Inbound & TaxiOut + Outbound represent the time it took the taxi to go get the passenger, so the car was empty for this leg of the taxi trip. Taxis to the airport for outbound flights have the highest density around 5 minutes and the distribution is extremely tight compared to all other distributions. 

```{r,warning=FALSE,cache=TRUE}
air_taxi %>%
  ggplot()+
  geom_smooth(aes(Month,TaxiTime))+
  facet_grid(Taxi~to_from)+
  scale_x_continuous(breaks = c(1:12))

```

Breaking Taxi times out in the same way as the previous one, but this time plotted against month of the year, seeing if there is a notable difference. There appears to be perhaps a little seasonality to this data (especially for TaxiOut Inbound flights), but not of particularly large magnitude.

```{r,warning=FALSE,cache=TRUE}
air_taxi %>%
  #filter((Origin %in% c("AUS"))) %>%
  filter(!is.na(TaxiTime)) %>%
  ggplot()+
  geom_jitter(aes(WeatherDelay,TaxiTime,group=Taxi,color=Taxi),alpha=.1)+
  geom_smooth(aes(WeatherDelay,TaxiTime,group=Taxi,color=Taxi),method = "lm")

```

Plotting the influence of weather delays on Taxi times, we see that weather does play a small but important role in the determining ride times. 

```{r,warning=FALSE,cache=TRUE}
air_taxi %>%
  #filter((Origin %in% c("AUS"))) %>%
  filter(!is.na(TaxiTime)) %>%
  ggplot()+
  geom_jitter(aes(DepTime,TaxiTime,group=Taxi,color=Taxi),alpha=.01)+
  geom_smooth(aes(DepTime,TaxiTime,group=Taxi,color=Taxi))+
  facet_grid(to_from~.)+
  ylim(0,50)

air_taxi %>%
  #filter((Origin %in% c("AUS"))) %>%
  filter(!is.na(TaxiTime)) %>%
  ggplot()+
  geom_jitter(aes(ArrTime,TaxiTime,group=Taxi,color=Taxi),alpha=.01)+
  geom_smooth(aes(ArrTime,TaxiTime,group=Taxi,color=Taxi))+
  facet_grid(to_from~.)+
  ylim(0,50)
```

This chart shows average TaxiIn and TaxiOut time by time of day.

Though both are extremely varied, there is a clear pattern in that TaxiIn is consistently lower than TaxiOut, suggesting rides to the airport are shorter than rides from the airport regardless of TOD. For TaxiOut, there is a clear change based on time of day, with modest increases in ride times around noon and 6pm, coinciding with high-traffic periods.

Lower TaxiIn adds to the hypothesis that TaxiIn rides may be coming from downtown, while TaxiOut rides tend to come from a wider radius of the ATX region.

## Question 4

```{r,warning=FALSE,cache=TRUE,message=FALSE}
sclass <- read.csv("../../../../../GitHub/ECO395M/data/sclass.csv")

library("tidyverse")
library("caret") 
library("foreach")
library("modelr")
library("rsample")

set.seed(1000)

sc350 <- sclass %>%
  filter(trim == "350") %>%
  select(mileage,price)

sc65 <- sclass %>%
  filter(trim == "65 AMG") %>%
  select(mileage,price)

sc350_split =  initial_split(sc350, prop=0.6)
sc350_train = training(sc350_split)
sc350_test  = testing(sc350_split)

sc65_split =  initial_split(sc65, prop=0.6)
sc65_train = training(sc65_split)
sc65_test  = testing(sc65_split)

```

### Trim 350

```{r,warning=FALSE,cache=TRUE}

k <- c(1:20,30,35,40, 45,
       50, 60, 70, 80, 90, 100, 125, 150, 175)

rmse_350 <- c()

plot350 <- sc350_test %>%
  ggplot()+
  geom_point(aes(mileage,price))

for(i in c(1:length(k))){
k350 <- knnreg(data = sc350_train,price ~ mileage,k=k[i])
rmse_350 <- c(rmse_350,rmse(k350,data = sc350_test))
sc350_test$predval <- predict(k350,newdata = sc350_test)
plot350 <- plot350+
  geom_line(data=sc350_test,aes(mileage,predval),alpha=.2,size=.8)
}

plot350

rmse_350 <- data.frame(cbind(rmse_350,k))

ggplot()+
  geom_point(data=rmse_350,aes(k,rmse_350))

print(paste("The value of k with lowest RMSE is:",rmse_350[rmse_350$rmse_350 == min(rmse_350$rmse_350),"k"]))
k_350 <- rmse_350[rmse_350$rmse_350 == min(rmse_350$rmse_350),"k"]

```

### Trim 65 AMG

```{r,warning=FALSE,cache=TRUE}

k <- c(1:25,30,35,40, 45,
       50, 60, 70, 80, 90, 100)

rmse_65 <- c()

plot65 <- sc65_test %>%
  ggplot()+
  geom_point(aes(mileage,price))

for(i in c(1:length(k))){
k65 <- knnreg(data = sc65_train,price ~ mileage,k=k[i])
rmse_65 <- c(rmse_65,rmse(k65,data = sc65_test))
sc65_test$predval <- predict(k65,newdata = sc65_test)
plot65 <- plot65+
  geom_line(data=sc65_test,aes(mileage,predval),alpha=.2,size=.8)
}

plot65

rmse_65 <- data.frame(cbind(rmse_65,k))

ggplot()+
  geom_point(data=rmse_65,aes(k,rmse_65))+
  theme_minimal()

print(paste("The value of k with lowest RMSE is:",rmse_65[rmse_65$rmse_65 == min(rmse_65$rmse_65),"k"]))
k_65 <- rmse_65[rmse_65$rmse_65 == min(rmse_65$rmse_65),"k"]

```

### Final Plot

```{r,warning=FALSE}
library("viridis")  
library("tidyverse")  


model_plot <- sclass %>% 
  filter(trim %in% c("350","63 AMG")) %>%
  ggplot()+
  geom_point(aes(mileage,price,color=trim),alpha=.1)+
  scale_color_manual(values=viridis(c(5))[c(2,4)])

sc350_test$predval <- predict(knnreg(data = sc350_train,price ~ mileage,k=k_350),
                              newdata = sc350_test)
model_plot <- model_plot +
  geom_line(data=sc350_test,aes(mileage,predval),color=viridis(c(5))[2],size=1)


sc65_test$predval <- predict(knnreg(data = sc65_train,price ~ mileage,k=k_65),
                              newdata = sc65_test)
model_plot <- model_plot +
  geom_line(data=sc65_test,aes(mileage,predval),color=viridis(c(5))[4],size=1)

model_plot
```

350 yields lower optimal K values, likely because it has many more observations and a dataset that is more regularly distributed. The 65 AMG data has a weird gap in the price data near the x axis which is likely causing issues with estimation and driving up RMSE.
